Example Manuscript Template for a Data Analysis Project

### 2024-04-07 products/manuscript/Manuscript.qmd
Still messing around with format and regression stuff
Still messing around with format and regression stuff

Influenza is an acute respiratory virus caused by one of several strains of virus.  Seasonal influenza is typically caused by circulating "A" and "B" strains of the disease.  Infection is accompanied by typical respiratory symptoms.  These may include cough, sore throat, and nasal discharge. Symptoms can also include abrupt onset of high-grade fever, muscle ache, loss of appetite, headache, and malaise.  After a typical course, influenza can go on to affect other organs such as the lungs, brain, and heart. Severe infections may require hospitalization.  Influenza infection can be much more severe for children, elderly adults, patients with chronic illnesses, and immunocompromised hosts.  According to data provided by the World Health organization, seasonal respiratory illnesses are linked to over 650,000 deaths each year worldwide. In the United States the flu is responsible for approximately 42,000 deaths per year and ranks 13th most frequent cause of death. As such, influenza is the subject of extensive public health surveillance and mediation efforts.**_[@Krammer-2018; @Moghadami-2017]_**

@@

```
**_Figure 1:_** *World Flu Burden by Clade.* Figure generated at GISAID.org.

@@

```
**_Figure 2:_** *World Flu Burden by Year.*

The graph reveals the extent of the data available.  The graph also reveals interesting features of the data set that may not have been readily apparent from if viewing the data in a tabular format.  Since 1995 the number of seasonal flu cases has followed a pattern that peaked during the winter months in the Northern hemisphere.  Notable features of this graph include the large jump in reported cases in the year 2010 and the large gap in reported cases in the year 2021.  The jump in cases on or about 2010 is due to increased participation in surveillance efforts and improved diagnostic techniques.  Hand in hand with these factors was an increasing awareness of the issue and consequent increased rate of reporting. **_[@Perdue-2012]_** The gap in the 2021 flu season was likely due to the measures taken to curb the SARS-CoV-2 pandemic through the 2020-2021 flu season.  Basic measures such as maintaining proper social distance, hand washing, public wearing of masks to prevent aerosol droplet spread, and other efforts were extremely effective at preventing the spread of seasonal flu. **_[@Rubin-2021]_**  Never the less, the amount of data reported worldwide since 1995 is massive.  In fact, the data is so dense that none of the categories of flu selected are visible as separate itmes in the bar graph.  Further filtering is obviously required.  For example, we might want to look at the total flu cases compared to the number of total cases of the A strains and the total number of the B strains in countries such as China, Australia, Canada, and France. It is also a good idea to clean up the data by omitting any missing values.  **Figure 3** depicts the data set filtered for information from China.

@@

```
**_Figure 3:_** *World Flu Burden by Year for China Since Reporting Began.*

@@

```
**_Figure 4:_** *World Flu Burden by Year: China, 2023-2024 Flu Season.*

@@

**_Figure 5:_** *World Flu Burden by Region Over the 2023-2024 Flu Season.*

@@

```
**_Figure 6:_** *World Flu Burden by Region Over the 2023-2024 Flu Season.*  WHO Regions: WPR, AMR, and EUR

@@

```
**_Figure 7:_** *World Flu Burden by Region Over the 2023-2024 Flu Season.* WHO Regions: AFR, EMR, SEAR

By splitting the graphs a great deal more detail can be discerned.  In addition to the putative west to east pattern that appeared for the WPR, AMR, and EUR WHO regions, there is a similar pattern for the SEAR and EMR regions.  An easy interpretation of this data would point to the seasonal flu beginning in the southeast Asia WHO region progressing to the eastern Mediterranean WHO region, then progressing to the WPR, AMR, and the EUR WHO regions respectively.  Of course, it is impossible to draw any conclusions from this data set and a single season.  To make a solid determination the data set would need to track not just time and geographic information, but connect the actual flu strains as they move around the globe.  This can be done with genetic data, i.e. DNA or RNA sequence data, from the flu cases reported.  In fact this is exactly the approach that the Nextstrain software takes.  In addition to presenting strong evidence of the central question of this research.  Further, graphing the data as opposed to attempting to genrate a summary table is a far more effective way to explore the data set.  As can be seen from **Table 1** (below), a tabular format doesn't highlight the patterns in the data in the same explicit manner as the graphic exploration.

@@

@@

*Explain anything related to your statistical analyses.*

@@

>>>>

Influenza is an acute respiratory virus caused by one of several strains of virus. Seasonal influenza is typically caused by circulating "A" and "B" strains of the disease. Infection is accompanied by typical respiratory symptoms. These may include cough, sore throat, and nasal discharge. Symptoms can also include abrupt onset of high-grade fever, muscle ache, loss of appetite, headache, and malaise. After a typical course, influenza can go on to affect other organs such as the lungs, brain, and heart. Severe infections may require hospitalization. Influenza infection can be much more severe for children, elderly adults, patients with chronic illnesses, and immunocompromised hosts. According to data provided by the World Health organization, seasonal respiratory illnesses are linked to over 650,000 deaths each year worldwide. In the United States the flu is responsible for approximately 42,000 deaths per year and ranks 13th most frequent cause of death. As such, influenza is the subject of extensive public health surveillance and mediation efforts.[@Krammer-2018***; @Moghadami-2017***]

@@

***Figure 1:*** *World Flu Burden by Clade.* Figure generated at GISAID.org.

@@

***Figure 2:*** *World Flu Burden by Year.*

The graph reveals the extent of the data available. The graph also reveals interesting features of the data set that may not have been readily apparent from if viewing the data in a tabular format. Since 1995 the number of seasonal flu cases has followed a pattern that peaked during the winter months in the Northern hemisphere. Notable features of this graph include the large jump in reported cases in the year 2010 and the large gap in reported cases in the year 2021. The jump in cases on or about 2010 is due to increased participation in surveillance efforts and improved diagnostic techniques. Hand in hand with these factors was an increasing awareness of the issue and consequent increased rate of reporting. [@Perdue-2012] The gap in the 2021 flu season was likely due to the measures taken to curb the SARS-CoV-2 pandemic through the 2020-2021 flu season. Basic measures such as maintaining proper social distance, hand washing, public wearing of masks to prevent aerosol droplet spread, and other efforts were extremely effective at preventing the spread of seasonal flu. [@Rubin-2021] Never the less, the amount of data reported worldwide since 1995 is massive. In fact, the data is so dense that none of the categories of flu selected are visible as separate itmes in the bar graph. Further filtering is obviously required. For example, we might want to look at the total flu cases compared to the number of total cases of the A strains and the total number of the B strains in countries such as China, Australia, Canada, and France. It is also a good idea to clean up the data by omitting any missing values. **Figure 3** depicts the data set filtered for information from China.

@@

***Figure 3:*** *World Flu Burden by Year for China Since Reporting Began.*

@@

***Figure 4:*** *World Flu Burden by Year: China, 2023-2024 Flu Season.*

@@

***Figure 5:*** *World Flu Burden by Region Over the 2023-2024 Flu Season.*

@@

***Figure 6:*** *World Flu Burden by Region Over the 2023-2024 Flu Season.* WHO Regions: WPR, AMR, and EUR

@@

***Figure 7:*** *World Flu Burden by Region Over the 2023-2024 Flu Season.* WHO Regions: AFR, EMR, SEAR

By splitting the graphs a great deal more detail can be discerned. In addition to the putative west to east pattern that appeared for the WPR, AMR, and EUR WHO regions, there is a similar pattern for the SEAR and EMR regions. An easy interpretation of this data would point to the seasonal flu beginning in the southeast Asia WHO region progressing to the eastern Mediterranean WHO region, then progressing to the WPR, AMR, and the EUR WHO regions respectively. Of course, it is impossible to draw any conclusions from this data set and a single season. To make a solid determination the data set would need to track not just time and geographic information, but connect the actual flu strains as they move around the globe. This can be done with genetic data, i.e. DNA or RNA sequence data, from the flu cases reported. In fact this is exactly the approach that the Nextstrain software takes. In addition to presenting strong evidence of the central question of this research. Further, graphing the data as opposed to attempting to generate a summary table is a far more effective way to explore the data set. As can be seen from **Table 1** (below), a tabular format doesn't highlight the patterns in the data in the same explicit manner as the graphic exploration.

@@

***Table 1:*** *World Flu Burden by Region Over the 2023-2024 Flu Season.*

@@

A cursory look at the data presented in **Figure 6** (above) shows a putative dependence of the AMR region flu season on the flu season peak approximately 3 week prior in the WPR region.  Likewise the peaks in reported flu cases in the EUR region seem to correspond to the WPR maxima approximately 3 weeks prior followed by a second maxima at a 3 week delay from the AMR region maxima.  A simple multiple linear regression allowing the EUR data to be dependent data, while using offset data from the WPR and AMR regions as independent data should reveal a strong relationship.

@@

***Table 1.*** \*

### 2024-04-07 products/manuscript/workspace.qmd
Still messing around with format and regression stuff
Still messing around with format and regression stuff

@@

@@

*Explain anything related to your statistical analyses.*

>>>>

```{r}
library(skimr)

#Setting up the data.
data_location <- here::here("data","raw-data","VIW_FNT_CM.xlsx")
WHO_data_flu <- readxl::read_excel(data_location)

WHO_data_flu <- WHO_data_flu %>% 
  select(WHOREGION, MMWR_WEEKSTARTDATE, INF_ALL)

#summarize the data in flu tot.
WHO_data_sum <- WHO_data_flu %>%
  group_by(WHOREGION, MMWR_WEEKSTARTDATE) %>%
  summarize(TOT_INF_ALL = sum(INF_ALL))

# Filter the data set for the selected regions
selected_regions1 <- c("WPR", "AMR", "EUR")
WHO_data_reg <- subset(WHO_data_sum, WHOREGION %in% selected_regions1)

#Generate the data sets for 2017-2018, 2018-2019, and 2019-2020
Sum_data_2017 <- WHO_data_reg %>% filter(MMWR_WEEKSTARTDATE > '2017-05-01' & MMWR_WEEKSTARTDATE < '2018-02-28')
Sum_data_2017 <- Sum_data_2017 %>% filter(MMWR_WEEKSTARTDATE > '2017-05-01' & MMWR_WEEKSTARTDATE < '2018-02-28')
Sum_data_2017 <- na.omit(Sum_data_2017)

Sum_data_2018 <- WHO_data_reg %>% filter(MMWR_WEEKSTARTDATE > '2018-05-01' & MMWR_WEEKSTARTDATE < '2019-02-28')
Sum_data_2018 <- Sum_data_2018 %>% filter(MMWR_WEEKSTARTDATE > '2018-05-01' & MMWR_WEEKSTARTDATE < '2019-02-28')
Sum_data_2018 <- na.omit(Sum_data_2018)

Sum_data_2019 <- WHO_data_reg %>% filter(MMWR_WEEKSTARTDATE > '2019-05-01' & MMWR_WEEKSTARTDATE < '2020-02-28')
Sum_data_2019 <- Sum_data_2019 %>% filter(MMWR_WEEKSTARTDATE > '2019-05-01' & MMWR_WEEKSTARTDATE < '2020-02-28')
Sum_data_2019 <- na.omit(Sum_data_2019)

# Graphs of prior flu years
# Season 2017 - 2018
plot2017 <- ggplot(Sum_data_2017, aes(x=MMWR_WEEKSTARTDATE)) +
  geom_point(aes(y=TOT_INF_ALL, color = WHOREGION)) +
  geom_line(aes(y=TOT_INF_ALL, color = WHOREGION)) +
    labs(x = "DATE", y = "Total Reported Cases of FLU", title = "Scatterplot Flu Cases Per WHO Region for the 2017-2018 Flu Season") +
  theme_minimal()

plot2017save = here("assets", "plot2017.png")
ggsave(plot2017save, plot2017, width = 10, height = 6, dpi = 300)

#| label: fig-schematic
#| fig-cap: "Figure XX: World Flu Burden by Region Over the 2023-2024 Flu Season."
#| echo: FALSE
knitr::include_graphics(here("assets","plot2017.png"))

# Season 2018-2019
plot2018 <- ggplot(Sum_data_2018, aes(x=MMWR_WEEKSTARTDATE)) +
  geom_point(aes(y=TOT_INF_ALL, color = WHOREGION)) +
  geom_line(aes(y=TOT_INF_ALL, color = WHOREGION)) +
    labs(x = "DATE", y = "Total Reported Cases of FLU", title = "Scatterplot Flu Cases Per WHO Region for the 2018-2019 Flu Season") +
  theme_minimal()

plot2018save = here("assets", "plot2018.png")
ggsave(plot2018save, plot2018, width = 10, height = 6, dpi = 300)

#| label: fig-schematic
#| fig-cap: "Figure XX: World Flu Burden by Region Over the 2018-2019 Flu Season."
#| echo: FALSE
knitr::include_graphics(here("assets","plot2018.png"))

#Season 2019 - 2020
plot2019 <- ggplot(Sum_data_2019, aes(x=MMWR_WEEKSTARTDATE)) +
  geom_point(aes(y=TOT_INF_ALL, color = WHOREGION)) +
  geom_line(aes(y=TOT_INF_ALL, color = WHOREGION)) +
    labs(x = "DATE", y = "Total Reported Cases of FLU", title = "Scatterplot Flu Cases Per WHO Region for the 2019-2020 Flu Season") +
  theme_minimal()

plot2019save = here("assets", "plot2019.png")
ggsave(plot2019save, plot2019, width = 10, height = 6, dpi = 300)

#| label: fig-schematic
#| fig-cap: "Figure XX: World Flu Burden by Region Over the 2018-2019 Flu Season."
#| echo: FALSE
knitr::include_graphics(here("assets","plot2019.png"))

@@

library(gt)
flu_data_wide |> 
  gt()

@@

*Explain anything related to your statistical analyses.*
```{r, echo=FALSE, message=FALSE}

suppressPackageStartupMessages(library(parsnip))
suppressPackageStartupMessages(library(tune))
suppressPackageStartupMessages(library(recipes))
suppressPackageStartupMessages(library(workflows))
suppressPackageStartupMessages(library(yardstick))
suppressPackageStartupMessages(library(rsample))

data_location <- here::here("data","raw-data","DATE_REGION.xlsx")
date_region <- readxl::read_excel(data_location)

date_region2 <- date_region %>% select(AFR, AMR, EMR, EUR, SEAR, WPR)

# fit the linear models with AMR as outcome 
# model has ALL OTHER REGIONS as predictors
lin_mod <- linear_reg() %>% set_engine("lm")
linfit1 <- lin_mod %>% fit(AMR ~ ., data = date_region2)

#model 2 -- includes date variable
lin_mod <- linear_reg() %>% set_engine("lm")
linfit2 <- lin_mod %>% fit(AMR ~ ., data = date_region)

# Compute the RMSE and R squared for model 1
metrics_1 <- linfit1 %>% 
  predict(date_region2) %>% 
  bind_cols(date_region2) %>% 
  metrics(truth = AMR, estimate = .pred)

# Compute the RMSE and R squared for model 2
metrics_2 <- linfit2 %>% 
  predict(date_region) %>% 
  bind_cols(date_region) %>% 
  metrics(truth = AMR, estimate = .pred)

lmtable1 <- broom::tidy(linfit1)
print(lmtable1)
print(metrics_1)

save_data_location <- here::here("assets","lmtable1.txt")
write.table(lmtable1, file = save_data_location, sep = "\t", row.names = FALSE)

lmtable2 <- broom::tidy(linfit2)
print(lmtable2)
print(metrics_2)

save_data_location <- here::here("assets","lmtable2.txt")
write.table(lmtable2, file = save_data_location, sep = "\t", row.names = FALSE)

data_location <- here::here("data","raw-data","REGION_OFFSET.xlsx")
Region_offset <- readxl::read_excel(data_location)

print(Region_offset)

lmfitB1 <- lm(EUR ~ WPR + AMR + EMR, Region_offset) 
lmtableB1 <- broom::tidy(lmfitB1)
print(lmtableB1)

logfitB1 <- glm(EUR ~ WPR + AMR + EMR, data = Region_offset)
logtableB1 <- broom::tidy(logfitB1)
print(logtableB1)
